ODE and Diffusion Models of Tumor Growth

### 2024-03-12 paper.tex
2 pages in

\section{Sample Math Equation}
Let's try displaying a math equation using LaTeX:
\[
    e^{i \pi} + 1 = 0
\]
This is Euler's famous equation, relating the mathematical constants $e$, $i$, $\pi$, and $0$.

\section{Lists}
LaTeX also makes it easy to create lists:

\subsection{Ordered List}
\begin{enumerate}
    \item First item
    \item Second item
    \item Third item
\end{enumerate}

\subsection{Unordered List}
\begin{itemize}
    \item Apples
    \item Oranges
    \item Bananas
\end{itemize}

>>>>

\section{Hill's Paper}

Hill's paper, titled `Diffusion of Oxygen and Lactic Acid through Tissues', set the scene for most of the diffusion models
developed in the late 20th century to present. In the introduction of this paper, Hill states that the 
`diffusion of dissolved substances through cells and tissues is a determining factor in many vital processes'. 
The diffusion constant, $k$, is introduced, along with its usages. It is representative of the number of unit quantities
of a substance that diffuse throughout an area of 1 $cm^2$ in a minute. The diffusion constant is usually a small value
when related to systems that are outside of the body; for example, the diffusion constant would be very small when studying
the diffusion of a gas within a large room. However, when studying systems within the body, the diffusion constant can be quite large.
Hill gives the following example when explaining this:

\begin{tabular}{|c|c|c|}
    \hline
    Object & Diameter & Time to 90\% Saturation \\
    \hline
    Cylinder & 1 cm & 185 minutes \\
    Actual Nerve & 0.7 mm & 54 seconds \\
    Single Nerve Fiber & 7 $\mu$ (0.0007 cm) & 0.0054 seconds \\
    \hline
\end{tabular}
\vspace*{1cm}

As seen by this example, many processes of diffusion throughout the body are extremely rapid. However, at the time of Hill's
paper, another problem was evident: not many problems in diffusion could be completely solved using mathematical methods.
Even when evaluating these equations, the existence of a unique solution was only guaranteed with certain inital conditions.
The steady state also needed to have specific parameters in order for the result to be of any interest.
\subsection{Diffusion Within a Steady State}

The first problem that Hill dealt with was the case where oxygen was being diffused from a gaseous or liquid phase.
It was assumed that the oxygen was being used up by metabolic processes at some rate $a$ and that the concentration
of oxygen, $y_0$, is maintained constant at $x = 0$.

\vspace*{0.5cm}

Let $x$ be the distance from the constant source of oxygen to any point within the tissue. Given the assumptions above,
the rate of diffusion across any unit area will be $ J = -k \* \frac{dy}{dx}$,\, where $J$ is the rate of diffusion, $k$ is the diffusion rate and the (-) 
represents that the diffusion is occuring in the opposite direction from the source. This is more commonly known as Fick's
first law of diffusion. To find the measure of this rate of accumulation, Fick's second law can be applied:
\begin{equation}
\frac{dJ}{dx} = k \* \frac{d^2y}{dx^2}
\end{equation}

Hill then combined these equations along with $a$, the rate of usage by the metabolic process, to come up with the first diffusion equation for
this model:

\begin{equation}
\frac{dy}{dt} + a = k \* \frac{d^2y}{dx^2}
\end{equation}
